---
title: "Prediction of the manner of exercise"
author: "Ivan Iakimov"
date: '08/05/2018'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Abstract

One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. A human activity recognition research has traditionally focused on discriminating between different activities, i.e. to predict "which" activity was performed at a specific point in time. Here we want to investigate "how (well)" an activity was performed by the wearer. We'll construct predictor to understand if a person do an exercise correctly or not.

## Data

Six young health participants were asked to perform one set of 10 repetitions of the Unilateral Dumbbell Biceps Curl in five different fashions: exactly according to the specification and with various types of errors. The dataset contains data from accelerometers on the belt, forearm, arm, and dumbell.

The description of the dataset can be found here:

[http://groupware.les.inf.puc-rio.br/har](http://groupware.les.inf.puc-rio.br/har)

Dataset at [https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv) contains information from accelerometers and also 'classe' variable, which describes if the participant executes the exercise correctly or with some sort of error.

## Goal

Our goal is to construct prediction function, which will take data from accelerometers on the belt, forearm, arm, and dumbell and provide information if the person executes the exercise correctly or which type of error she has.

## Setup

Here I will use:   

- R version 3.3.2 (2016-10-31)  
- Platform: x86_64-w64-mingw32/x64 (64-bit)  
- Running under: Windows 10 x64  
- Intel(R) Core(TM) i7-3630QM CPU @ 2.40 GHz  
- 16 GB of memory  
- NVidia NVS 5200M GPU  

I'll use the following packages:

```{r}
require(dplyr)
require(caret)
```

## Data analysis

Let's take a look at the dataset:

```{r}
if(!file.exists("pml-training.csv"))
{
    download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", 
                  destfile = "pml-training.csv");
}

data <- read.csv("pml-training.csv")

data <- tbl_df(data)

dim(data)
```

Let's clean the data a little bit. First of all, I'll remove time and user information:

```{r}
data <- select(data, -(2:7))
```

This dataset contains a lot of columns with many NA values. 

```{r}
summary(data$max_roll_belt)
```

I'll remove such columns from the dataset:

```{r}
hasNA <- function(x) { sum(is.na(x)) > 0 }

columns.with.na <- lapply(data, hasNA)

names.of.columns.without.na <- names(data)[!as.logical(columns.with.na)]

data <- select(data, names.of.columns.without.na)

rm("names.of.columns.without.na", "columns.with.na", "hasNA")
```

Now there are also columns containing data in character format. They also contain a lot of empty values:

```{r}
sum(data$kurtosis_roll_belt == "")
```

So I'll remove them too:

```{r}
factor.columns <- lapply(data, class)

names.of.non.factor.columns.with.class <- c("classe", names(data)[factor.columns != "factor"])

data <- select(data, names.of.non.factor.columns.with.class)

rm("factor.columns", "names.of.non.factor.columns.with.class")
```

## Prediction model creation

Now we can build a prediction model on the cleaned dataset. I'll use random forest:

```{r, cache=TRUE}
set.seed(45273)

inTrain <- createDataPartition(y = data$classe, p = 0.6, list = FALSE)

training <- data[inTrain,]
testing <- data[-inTrain,]

rm("inTrain")

model.fit <- train(classe ~ ., data = training, method = "rf")
```

Now we can estimate accuracy of our model. On the training dataset:

```{r}
pred.train <- predict(model.fit, training)

cm <- confusionMatrix(pred.train, training$classe)
```

accuracy equals to `r cm$overall[["Accuracy"]]` with confidence interval (`r cm$overall[["AccuracyLower"]]`; `r cm$overall[["AccuracyUpper"]]`)

On the testing dataset:

```{r}
pred.test <- predict(model.fit, testing)

cm <- confusionMatrix(pred.test, testing$classe)
```

accuracy equals to `r cm$overall[["Accuracy"]]` with confidence interval (`r cm$overall[["AccuracyLower"]]`; `r cm$overall[["AccuracyUpper"]]`)

As you can see, the model is pretty accurate.

## Prediction for real data

We can test our prediction model on some real data from [https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv](https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv).

```{r}
if(!file.exists("pml-testing.csv"))
{
    download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", 
                  destfile = "pml-testing.csv");
}

validate <- read.csv("pml-testing.csv")

validate <- tbl_df(validate)

pred.valid <- predict(model.fit, validate)

pred.valid
```

In this testing dataset all participant have class `r as.character(pred.valid[1])` of the exercise execution.

## Conclusion

In this work we created prediction model for recognizing class of exercise execution using data from different accelerometers. The model is based on random forest and has good accuracy.

We predicted the class of exercise execution for a test dataset and all participants have class `r as.character(pred.valid[1])`.